---
title: "Assignment 1"
author: "Hongyu He (2632195) & Bruno Hoevelaken (2645065)"
date: "Group CS 6"
output: pdf_document
---
## Theoretical Exercises
### Exercise 1.1
1. (Section 1.2 E10)
The sampling method used in this scenario is **flawed**. *The book clearly states* (p7) **that internet polls**, in which people online can decide whether to respond, **are by their very nature seriously flawed**, as they result in **voluntary response samples**. This also means **nonresponses** might have occuered. This would  result in **missing data** regarding individuals who were unwilling to participate in the online poll. Due to the missing data, we are unable extrapolate any meaningful conclusions out of this sample, regarding the wider populous. For a voluntary response sample, we can draw valid conclusions only about the specific group of people who chose to participate. From a statistical viewpoint, such a method is fundamentally flawed and should not be used for making general statements about a large population.

2. (Section 1.2 E12)
In this case, the 1018 samples used by Gallup pollsters are *simple random samples* so the sampling method appears to be **sound**.

3. (Section 1.2 E26)
We believe that the sampling method is **flawed**. Same as "Section 1.1 E10", the online poll uses *voluntary response samples* (or *self-selected samples*). Specifically, the *Internet poll* method is used in this survey. According to the book: *"By their very nature, all are seriously flawed because we should not make conclusions about a population on the basis of such a biased sample."*. 
This exercise presents a scenario which suffers from the *same issue as exercise 10*, as it relies on an **internet poll**. Another issue is the fact that it is unclear what the other 59% of users said.
The result, therefore, is most likely to be *incorrect*.

### Exercise 1.2
1. (Section 1.3 E22) 
Depths (km) of earthquakes are at the **ratio** level of measurement because the depths (km) of earthquakes measured relative to the earth's crust(s), which means there is a meaningful **natural zero starting point**.

2. (Section 1.3 E32)
The lineup numbers exist as distinguishing features, each corresponding to a player. This means that they fall under the **nominal** level of measurement. So *calculating their mean is pointless*.

### Exercise 1.3
1. (Section 1.4 E6)
This should be an **experiment** since the researchers first applied some treatment to the subjects and then observed the effects on the experimental units.

2. (Section 1.4 E12)
The researcher used **systematic sampling** because every 5 meters, a sample was selected. 

3. (Section 1.4 E18)
In this case, the sampling method used by ABC News should be **cluster sampling** where the voters were (implicitly) divide into clusters (stations) and then those stations were randomly selected. Finally, all members (voters) of the selected clusters are chosen.

### Exercise 1.4
#### a. 
- first off, the y-axis does not start from 0;
- the step size of the y-axis is too large, which makes hard reading;
- the bars are not in the same width.

#### b. 
1. Since the analysed data are at the nominal level of measurement, I think both **Pareto chart** and **pie chart** are good choices for illustrating the distribution of the food preference. In addition, the relative frequencies are needed to be calculated first for making a pie chart.

2. Due to the measured data are at the ratio level of measurement, I think both **histogram** and **boxplot** are illegible for this case.  

## R-exercises
### Exercise 1.5
#### a. 
```{r echo=FALSE, out.width = '50%', out.height = '40%'}
# a.
A_data = scan("sampleA.txt")

hist(A_data, breaks = 20, xlim = range(0, 10), axes = FALSE, col = "darksalmon", main = "Histogram of sampleA")
xlabel <- seq(0, 9, by = .5)
axis(1, at = xlabel)
ylabel <- seq(0, 15, by = 1)
axis(2, at = ylabel)
boxplot(A_data, outlier.tagging = TRUE, col = "darksalmon", main = "Boxplot of sampleA")

```

### b. Numerical summaries for *sampleA*

|Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max. | Variance | Standard deviaton|
|:-----:|:--------------:|:--------:|:------:|:--------------:|:-----:|:---:|:---:|
|0.1087|  1.2170 | 2.4740 |  2.8300|  4.0360 | 8.5560 | 3.662593|1.91379|

```{r echo=FALSE, out.width = '50%', out.height = '40%'}

library(knitr)
#title = c("Min.", "1st Quartile", "Median", "Mean", "3rd Quartile", "Max.")
#describe = cbind(title, summary(A_data))
#kable(describe)
 #summary(A_data)
#print("Variance:")
#var(A_data)
#print("Standard deviaton:")
#sd(A_data)
```

### c.
- As we can see from the two diagrams, the data is **right-skewed**;
- Since the data is quite right-skewed, the **accumulation** is relatively low;
- Clearly, this data is **unimodal (mode/peak is around 0.8)** and not **symmetric**;
- Based on the **median** value, the **location** of the data is around 2; 
- The **range** of the data is 8.5560 - 0.1087 = 8.4472;
- There is one **extreme** value (**outlier**) which is 8.5560;
- Comparing the mean and standard deviation value, the **variation** of the data is quite big.

### d. Plots and Numerical summaries for *sampleB*

```{r echo=FALSE, out.width = '50%', out.height = '40%'}
B_data = scan("sampleB.txt")

hist(B_data, breaks = 20, axes = FALSE, col = "bisque", main = "Histogram of sampleB")
xlabel <- seq(0, 6, by = .5)
axis(1, at = xlabel)
ylabel <- seq(0, 25, by = 1)
axis(2, at = ylabel)
boxplot(B_data, outlier.tagging = TRUE, col = "bisque", main = "Boxplot of sampleB")

#summary(B_data)
#var(B_data)
#sd(B_data)
```

|Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max. | Variance | Standard deviaton|
|:-----:|:--------------:|:--------:|:------:|:--------------:|:-----:|:---:|:---:|
|0.01543 | 2.01700 | 2.82400 | 2.92900 | 3.87400 | 5.97900 | 1.7051 | 1.305795|

- As we can see from the two diagrams, the data is **roughly symmetric**;
- Compared to `A_data`, the **accumulation** of this data set is faster;
- This data is **unimodal** (mode/peak is around 2.2);
- Based on the **median** value, the location of the data is around 2.5; 
- The **range** of the data is 5.96357;
- There is no **extreme** value (**outlier**);
- Comparing the mean and (sample) standard deviation value, the **variation** of the data is relatively small.

### e. 
  According to the previous analyses on the two data sets, none of the characteristics (shape, accumulation, mode, location, range, outliers, variation) of them are alike. Therefore, I strongly convinced that they are not originated from the same population distribution.


## Exercise 1.6
### (1)
```{r echo=FALSE, out.width = '70%', out.height = '40%'}
par(mfrow=c(1,3))
mileage = source("mileage.txt")
attach(mileage$value)

y_label = "Miles per gallon"
boxplot(mpg1[cyl1==4], mpg2[cyl1==4], ylim = range(10:40), names = c("Type1", "Type2"), ylab = y_label, col = "chartreuse2", main = "Cars with 4 cylinders")
boxplot(mpg1[cyl1==6], mpg2[cyl1==6], ylim = range(10:40), names = c("Type1", "Type2"), ylab = y_label, col = "coral1", main = "Cars with 6 cylinders")
boxplot(mpg1[cyl1==8], mpg2[cyl1==8], ylim = range(10:40), names = c("Type1", "Type2"), ylab = y_label, col = "gold1", main = "Cars with 8 cylinders")

```

- 4 Cylinders:

|Type|Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max. |
|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|
|Type1| 21.50 | 27.45 | 30.50 | 30.02 | 32.70 |
|Type2| 21.40 | 22.80 | 26.00 | 26.66 | 30.40 |

- 6 Cylinders:

|Type|Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max. |
|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|
| 15.50        | 17.75  | 22.00        | 24.25 | 29.65 | 37.30        |
| 17.80        | 18.65  | 19.70        | 19.74 | 21.00 | 21.40        |

- 8 Cylinders:

|Type|Min. | 1st Quartile | Median | Mean | 3rd Quartile | Max. |
|:-----:|:-----:|:-----:|:-----:|:-----:|:-----:|
| 15.50        | 16.80  | 17.30        | 17.42 | 18.27 | 19.20        |
| 10.40        | 14.40  | 15.20        | 15.10 | 16.25 | 19.20 |

The 3 plots above compare the fuel usage of car type 1 & 2 under the same number of cylinders. As they illustrate, when the numbers of cylinders are 4 and 6, the type 2 cars are generally more fuel-efficient than type 1 cars, whilst for 8-cylinder cars, type 1 is more fuel-efficient than type 2.

### (2)
```{r echo=FALSE, out.width = '70%', out.height = '40%'}
## (2)
par(mfrow=c(1,2))
x_names <- seq(4, 8, by = 2)
x_label = "Number of cylinders"

boxplot(mpg1[cyl1==4], mpg1[cyl1==6], mpg1[cyl1==8], ylim = range(10:40), names = x_names, xlab = x_label, ylab = y_label, col = "deepskyblue", main = "Car type 1")
boxplot(mpg2[cyl2==4], mpg2[cyl2==6], mpg2[cyl2==8], ylim = range(10:40), names = x_names, xlab = x_label, ylab = y_label, col = "firebrick1", main = "Car type 2")

```

No, I think there indeed a risk to directly compare the data of both types of cars when including also the cars with more cylinders. As we can see from the two boxplots above, for both type 1 & 2, the cars with more cylinders are always more fuel-efficient than the ones with fewer cylinders.


## ***Appendix***
```r
# Exercise 1.5
# a.
A_data = scan("sampleA.txt")

hist(A_data, breaks = 20, xlim = range(0, 10), axes = FALSE, col = "darksalmon", main = "Histogram of sampleA")
xlabel <- seq(0, 9, by = .5)
axis(1, at = xlabel)
ylabel <- seq(0, 15, by = 1)
axis(2, at = ylabel)
boxplot(A_data, outlier.tagging = TRUE, col = "darksalmon", main = "Boxplot of sampleA")

# b.
summary(A_data)
var(A_data)
sd(A_data)

# d.
B_data = scan("sampleB.txt")

hist(B_data, breaks = 20, axes = FALSE, col = "bisque", main = "Histogram of sampleB")
xlabel <- seq(0, 6, by = .5)
axis(1, at = xlabel)
ylabel <- seq(0, 25, by = 1)
axis(2, at = ylabel)
boxplot(B_data, outlier.tagging = TRUE, col = "bisque", main = "Boxplot of sampleB")

summary(B_data)
var(B_data)
sd(B_data)

# Exercise 1.6
# (1)
par(mfrow=c(1,3))
mileage = source("mileage.txt")

y_label = "Miles per gallon"
boxplot(mpg1[cyl1==4], mpg2[cyl1==4], ylim = range(10:40), names = c("Type1", "Type2"), ylab = y_label, col = "chartreuse2", main = "Cars with 4 cylinders")
boxplot(mpg1[cyl1==6], mpg2[cyl1==6], ylim = range(10:40), names = c("Type1", "Type2"), ylab = y_label, col = "coral1", main = "Cars with 6 cylinders")
boxplot(mpg1[cyl1==8], mpg2[cyl1==8], ylim = range(10:40), names = c("Type1", "Type2"), ylab = y_label, col = "gold1", main = "Cars with 8 cylinders")

summary(mpg1[cyl1==4])
summary(mpg2[cyl2==4])

summary(mpg1[cyl2==6])
summary(mpg2[cyl2==6])

summary(mpg1[cyl1==8])
summary(mpg2[cyl2==8])

# (2)
par(mfrow=c(1,2))
x_names <- seq(4, 8, by = 2)
x_label = "Number of cylinders"

boxplot(mpg1[cyl1==4], mpg1[cyl1==6], mpg1[cyl1==8], ylim = range(10:40), names = x_names, xlab = x_label, ylab = y_label, col = "deepskyblue", main = "Car type 1")
boxplot(mpg2[cyl2==4], mpg2[cyl2==6], mpg2[cyl2==8], ylim = range(10:40), names = x_names, xlab = x_label, ylab = y_label, col = "firebrick1", main = "Car type 2")
```
