---
title: "Assignment 3"
author: "Hongyu He (2632195) & Bruno Hoevelaken (2645065)"
date: "Group CS 6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# & Theoratical exercises

## Exercise 3.1 (Section 8.4, E16)

Based on the setting of this problem, the two samples are **two dependent samples** (Matched Pairs). Therefore, their **population deviations** $\sigma_1$ and $\sigma_2$ are considered to be the same.

### 1) The hypotheses in terms of the population parameter of interest:

- H~0~: $\mu$~d~ = 0; (original claim)
- H~a~: $\mu$~d~ $\neq$ 0;

### 2) The significance level: 

- $\alpha$ = 5% (two-sided)

### 3) The test statistic and its distribution under the null hypothesis:

According to the problem, the standard deviation of the population is **unknown**. Therefore, in this test, we replace it by the estimator S~d~. The test statistic (two-tailed) has t-distribution with (n-1 = 11) degrees of freedom (`fd`).

### 4) The observed value of the test statistic (the observed score);

Under the assumption that the null hypothesis is true:

$$ 
t =\frac{\bar{d}-\mu_{d}}{\frac{s_{d}}{\sqrt{n}}} = \frac{ (\bar{X}_1 - \bar{X}_2) - 0 }{ S_d/ \sqrt{n} } = -0.984
$$

### 5) The P-value or the critical values
  Since we assume technology is not available now, so we hereby only calculated the **critical values**:
  
- `df` = n - 1 = 11
  
- $-$ 2.718 < $-$ t~11,0.025~ < $-$ 2.201 (Tech: $-$ t~11,0.025~ = $-$ 2.200985)

- $+$ 2.201 < $+$ t~11,0.025~ < $+$ 2.718 (Tech: $+$ t~11,0.025~ = $+$ 2.200985)

### 6) Whether or not the null hypothesis is rejected and why.
  Since t > $-$ t~11,0.025~ (Tech: > $-$ 2.200985), the test statistics, therefore, is not in the critical region.
  
Thus, 

- we fail to reject the null hypothesis;
- there is not enough evidence warrant the rejection of the claim that males aged 12-16 correctly report their heights.


## Exercise 3.2 (Section 6.2, E32)

According to the problem, we know that 

- E = 0.03
- Confidence level = 99%
- $\hat{p}$ = 15%

Therefore, 

- $\alpha$ /2 = (1-99%) / 2 = 0.005
- $\hat{q}$ = 1 - 15% = 85%

and so

- $Z_\alpha/2$ = 2.575

Thus, the sample size is

$$
\begin{aligned} n &=\frac{\left[z_{\alpha / 2}\right]^{2} \hat{p} \hat{q}}{E^{2}}=\frac{[2.575]^{2}(0.15)(0.85)}{0.03^{2}} \\ &=939.34 \geq 940 \end{aligned}
$$

Therefore, the **sample size** n should be greater or equal to **940**.

# & R-exercises

## Exercise 3.3

### a)

```{r echo=FALSE, out.width = '20%', out.height = '20%'}
load("Assign3.RData") # move to appendix
```
```{r echo=TRUE}
t.test(Alice, Bob, mu = 0, alternative = "two.sided", paired=TRUE, conf.level = 0.95)
```
According to the above results:

- An **estimate** for the difference of mean working time per evening of Alice and Bob: **0.2689**
- A 95% confidence interval for the difference of mean working time per evening of Alice and Bob: **[0.1142398, 0.4235934]**

### b)

#### b1) R command and results:

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
t.test(Alice, Bob, mu = 0, alternative = "two.sided", paired = TRUE, conf.level = 0.90)
```

#### b2) The hypotheses in terms of the population parameter of interest:

- H~0~: $\mu$~d~ = 0; (original claim)
- H~a~: $\mu$~d~ $\neq$ 0;

#### b3) The significance level: 

- $\alpha$ = 10%

#### b4) The test statistic and its distribution under the null hypothesis:

- two-tailed test;
- two dependent samples (Matched Pairs);
- the test statistic is under a t-distribution.

#### b5) The observed value of the test statistic (the observed score);

Under the assumption that the null hypothesis is true, the test statistic is 3.4938.

#### b6) The P-value

- P-value = 0.00102

#### b7) Conclusions

According to the results of above the t-test command, we know that the **p-value** is **0.001**. 

Since the **p-value=0.001 < $\alpha$=0.01**, our conclusions are as follows:

- reject the null-hypothesis;
- there is sufficient evidence to warrant rejection of the claim that on average Alice and Bob both work the same amount of time.

### c)

#### c1) R command and results:

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
t.test(Alice, Bob, mu = 0, alternative = "greater", paired = TRUE, conf.level = 0.90)
```

#### c2) The hypotheses in terms of the population parameter of interest:

- H~0~: $\mu$~d~ = 0; 
- H~a~: $\mu$~d~ $>$ 0; (original claim)

#### c3) The significance level: 

- $\alpha$ = 10%

#### c4) The test statistic and its distribution under the null hypothesis:

- right-tailed test;
- two dependent samples (Matched Pairs);
- the test statistic is under a t-distribution.

#### c5) The observed value of the test statistic (the observed score);

Under the assumption that the null hypothesis is true, the test statistic is 3.4938.

#### c6) The P-value

- P-value = 0.0005101

#### c7) Conclusions

According to the results of the above t-test command, we know that the **p-value** is **0.0005**. 

Since the **p-value=0.0005 < $\alpha$=0.01**, our conclusions are as follows:

- reject the null-hypothesis;
- there is sufficient evidence to support the claim that Alice works, on average, much more than Bob.


### d)

#### (1) Changes

Yes, because this time the data contained in *Assign3.RData* **is no longer a dependent (matched pairs) sample** in the sense that the data of `Alice` and `Bob` correspond to different workdays. 

Consequently, as mentioned in the problem, the $\sigma_1$ of sample vector `Alice` and the $\sigma_2$ of sample vector `Bob` is different. 

#### (2) Rre-test

##### d1) R command and results:

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
t.test(Alice, Bob, mu = 0, alternative = "two.sided", 
       paired = FALSE, var.equal = FALSE, conf.level = 0.90)
```


##### d2) The hypotheses in terms of the population parameter of interest:

- H~0~: $\mu$~d~ = 0; (original claim)
- H~a~: $\mu$~d~ $\neq$ 0;

##### d3) The significance level: 

- $\alpha$ = 10%

##### d4) The test statistic and its distribution under the null hypothesis:

- two-tailed test;
- two independent samples;
- $\sigma_1 \ and\ \sigma_2$ unknown and we assume that $\sigma_1 \neq\sigma_2$;
- the test statistic is under a t-distribution.

##### d5) The observed value of the test statistic (the observed score);

Under the assumption that the null hypothesis is true, the test statistic is 3.3647.

##### d6) The P-value

- P-value = 0.001256;

##### d7) Conclusions

According to the results of the above t-test command with a new setting, we know that the p-value is 0.001. 

Since the **p-value=0.001 < $\alpha$=0.01**, our conclusions are the same as a):

- reject the null-hypothesis;
- there is sufficient evidence to warrant rejection of the claim that on average Alice and Bob both work the same amount of time.


## Exercise 3.4

### a)

A **point estimate** that we obtained for the difference in the proportion of evenings that Alice and Bob have worked more than 4 hours is **0.36**.


### b)

#### b1) R command and results:

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
prop.test(x = c(length(Alice[Alice>4]), length(Bob[Bob>4])), 
          n = c(50, 50), alternative = "greater", conf.level = 0.95)
```

#### b2) The hypotheses in terms of the population parameter of interest:

- H~0~: $\mu$~d~ = 0; 
- H~a~: $\mu$~d~ $>$ 0; (original claim)

#### b3) The significance level: 

- $\alpha$ = 5%

#### b4) The test statistic and its distribution under the null hypothesis:

- right-tailed;
- two independent samples;
- the test statistic is under a t-distribution.

#### b5) The observed value of the test statistic (the observed score);

Under the assumption that the null hypothesis is true:

- the pooled sample proportion: $\bar{p}=\frac{x_{1}+x_{2}}{n_{1}+n_{2}} = 0.62$

- the test statistic:
$z=\frac{\left(\hat{p}_{1}-\hat{p}_{2}\right)-\left(p_{1}-p_{2}\right)}{\sqrt{\frac{\bar{p} \bar{q}}{n_{1}}+\frac{\bar{p} \bar{q}}{n_{2}}}} = 3.708$.

#### b6) The P-value

- P-value = 0.0002306

#### b7) Conclusions

The results of the above two proportions test results in a **p-value** of **0.0002** which is **smaller than** the significance level **$\alpha$ = 0.05**.

Therefore, our conclusions are as follows:

- We reject the null hypothesis;
- There is sufficient evidence to support the claim that the proportion of evenings on which she worked more than 4 hours is larger than the proportion of evenings on which Bob worked more than 4 hours.


# *Appendix*
```r
#E3.1
qt(0.05/2, 11)
abs(qt(0.05/2, 11))

# E3.3
load("Assign3.RData")

# a)
alice_propotion = mean(Alice > 4)
bob_propotion = mean(Bob > 4)

point_estimate = alice_propotion - bob_propotion

# b)
pooled_p = (length(Alice[Alice>4]) + length(Bob[Bob>4])) / (50*2)

test_statistic = (alice_propotion - bob_propotion) / sqrt(2*pooled_p*(1-pooled_p) / 50)
```









