---
title: "Assignment 3"
author: "Hongyu He (2632195) & Bruno Hoevelaken (2645065)"
date: "Group CS 6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# & Theoratical exercises

## Exercise 3.1 (Section 8.4, E16)

Based on the setting of this problem, the two samples are two dependent samples (Matched Pairs). Therefore, their population deviations $\sigma_1$ and $\sigma_2$ are considered to be the same.

### 1) The hypotheses in terms of the population parameter of interest:

- H~0~: $\mu$~d~ = 0; (original claim)
- H~a~: $\mu$~d~ $\neq$ 0;

### 2) The significance level: 

- $\alpha$ = 5%

### 3) The test statistic and its distribution under the null hypothesis:

According to the problem, the standard deviation of the population is unknown. Therefore, in this test, we replace it by the estimator S~d~. The test statistics has t-distribution with (n-1 = 11) degrees of freedom.

### 4) The observed value of the test statistic (the observed score);

The observed value of test statistics is:

$$ 
t = \frac{ (\bar{X}_1 - \bar{X}_2) - 0 }{ S_d/ \sqrt{n} } = -0.984
$$

### 5) The P-value or the critical values;
  Since we assume technology is not available now, so we only calculated the critical values -t~11,0.025~ =  -2.201 and t~11,0.025~ =  2.201

### 6) Whether or not the null hypothesis is rejected and why.
  Since t > -t~11,0.025~, the test statistics, therefore, is not in the critical region.
  
Thus, 

- we fail to reject the null hypothesis;
- there is not enough evidence warrant the rejection of the claim that males aged 12-16 correctly report their heights.


## Exercise 3.2 (Section 6.2, E32)

According to the problem, we know that 
- E = 0.03
- Confidence level = 99%
- $\hat{p}$ = 15%

Therefore, 

- $\alpha$ /2 = (1-99%) / 2 = 0.005
- $\hat{q}$ = 1 - 15% = 85%

and so

- $Z_\alpha/2$ = 0.5199

Thus, the sample size is

$$
\begin{aligned} n &=\frac{\left[z_{\alpha / 2}\right]^{2} \hat{p} \hat{q}}{E^{2}}=\frac{[0.5199]^{2}(0.15)(0.85)}{0.03^{2}} \\ &=38.291=39 \end{aligned}
$$


# & R-exercises

## Exercise 3.3

### a)

```{r echo=FALSE, out.width = '20%', out.height = '20%'}
load("Assign3.RData") # move to appendix
```
```{r echo=TRUE}
t.test(Alice, Bob, mu = 0, alternative = "two.sided", paired=TRUE, conf.level = 0.95)
```
According to the above results:

- An estimate for the difference of mean working time per evening of Alice and Bob: 0.2689
- A 95% confidence interval for the difference of mean working
time per evening of Alice and Bob: [0.1142398, 0.4235934]

### b)

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
t.test(Alice, Bob, mu = 0, alternative = "two.sided", paired = TRUE, conf.level = 0.90)
```

According to the results of above the t-test command, we know that the p-value is 0.001. Since the p-value=0.001 < $\alpha$=0.01, our conclusions are as follows:

- reject the null-hypothesis;
- there is sufficient evidence to warrant rejection of the claim that on average Alice and Bob both work the same amount of time.

### c)

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
t.test(Alice, Bob, mu = 0, alternative = "greater", paired = TRUE, conf.level = 0.90)
```

According to the results of the above t-test command, we know that the p-value is 0.0005. Since the p-value=0.0005 < $\alpha$=0.01, our conclusions are as follows:

- reject the null-hypothesis;
- there is sufficient evidence to support the claim that Alice works, on average, much more than Bob.

### d)

#### (1) Changes

Yes, because this time the data contained in *Assign3.RData* is no longer a dependent (matched pairs) sample in the sense that the data of Alice and Bob correspond to different workdays. 
Consequently, as mentioned in the problem, the $\sigma_1$ of sample vector `Alice` and the $\sigma_2$ of sample vector `Bob` is different. 

#### (2) Rre-test

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
t.test(Alice, Bob, mu = 0, alternative = "two.sided", 
       paired = FALSE, var.equal = FALSE, conf.level = 0.90)
```

According to the results of the above t-test command with a new setting, we know that the p-value is 0.001. Since the p-value=0.001 < $\alpha$=0.01, our conclusions are the same as a):

- reject the null-hypothesis;
- there is sufficient evidence to warrant rejection of the claim that on average Alice and Bob both work the same amount of time.

## Exercise 3.4

### a)

A **point estimate** that we obtained for the difference in the proportion of evenings that Alice and Bob have worked more than 4 hours is **0.36**.


### b)

```{r echo=TRUE, out.width = '20%', out.height = '20%'}
prop.test(x = c(length(Alice[Alice>4]), length(Bob[Bob>4])), 
          n = c(50, 50), alternative = "greater", conf.level = 0.95)
```

The results of the above two proportions test results in a p-value of 0.0002 which is smaller than the significance level $\alpha$ = 0.05.

Therefore, our conclusions are as follows:

- We reject the null hypothesis;
- There is sufficient evidence to support the claim that the proportion of evenings on which she worked more than 4 hours is larger than the proportion of evenings on which Bob worked more than 4 hours.




```{r echo=FALSE, out.width = '60%', out.height = '50%'}

```
```{r echo=FALSE, out.width = '60%', out.height = '50%'}

```


# *Appendix*
```r
# E3.3
load("Assign3.RData")
```









